---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# ggsomewhere

<!-- badges: start -->
<!-- badges: end -->

The goal of ggsomewhere (and it's github repository) is to provide tools for building spatial ggplot2 extensions.

The approach makes use of ggplot2's sf plotting capabilities, the sf2stat package to prepare reference data for use inside a stat_* or geom_* layer, and the readme2pkg package functions for managing functions from a readme and working with template functions included in the readme.

## Fully worked example...

How you'd use sf2stat to build functionality with scope, region type, and location name.  When using the code templates for this north carolina county example, we'll be replacing 'scope', 'region', and 'locations' as follows

- 'scope' -> 'brasil'
- 'region' -> 'county'
- 'locations' -> a vector of region names

Let's see how we might recreate the functionality in the ggnorthcarolina package using some templates in this readme.

In the example, the scope of the package is 'northcarolina'.  The region of interest is 'county', and the location names that we are using are the county names.

##

```{r, eval = F}
devtools::create(".")
```

## Step 00. prep reference data

### Create 

```{r, eval = F}
usethis::use_data_raw()
```

## Preparation

The reference data should have just the id columns and the geometry.  Then it should be named similar to this: 

geo_reference_brasil_

Where 'northcarolina' is is the character string that will replace 'scope' when using the code templates, and 'county' is the character string that will replace region in the 


```{r geo_reference_brasil_state, eval = T}
br_state <- geobr::read_state(simplified = T)
br_muni <- geobr::read_municipality()


geo_reference_brasil_state <- br_state |>
  dplyr::select(state_name = name_state, state_code = code_state, state_abbrev = abbrev_state, geometry = geom) |>
  sf2stat:::sf_df_prep_for_stat(id_col_name = "state_name")

geo_reference_brasil_muni <- br_muni |>
  dplyr::select(muni_name = name_muni, muni_code = code_muni, state_abbrev = abbrev_state, geometry = geom) |>
  sf2stat:::sf_df_prep_for_stat(id_col_name = "muni_name")

usethis::use_data(geo_reference_brasil_state, overwrite = T)
usethis::use_data(geo_reference_brasil_muni, overwrite = T)

```


```{r, eval = F}
readme2pkg::chunk_to_dir("geo_reference_brasil_state", 
                         dir = "data-raw/")
```

## Use template to create `stat_county()` functionality

```{r}
readme2pkg::chunk_variants_to_dir(chunk_name = "stat_region_template",
                                  file_name = "stat_state.R",
                                  replace1 = "scope",
                                  replacements1 = "brasil",
                                  replace2 = "region",
                                  replacements2 = "state")

readme2pkg::chunk_variants_to_dir(chunk_name = "stat_region_template",
                                  file_name = "stat_muni.R",
                                  replace1 = "scope",
                                  replacements1 = "brasil",
                                  replace2 = "region",
                                  replacements2 = "muni")

```


```{r stat_region_template, eval = F}
compute_panel_scope_region <- function(data, 
                                       scales, 
                                       keep_id = NULL, 
                                       drop_id = NULL, 
                                       stamp = FALSE){
  
  if(!stamp){data <- dplyr::inner_join(data, geo_reference_scope_region)}
  if( stamp){data <- geo_reference_scope_region }
  
  if(!is.null(keep_id)){ data <- dplyr::filter(data, id_col %in% keep_id) }
  if(!is.null(drop_id)){ data <- dplyr::filter(data, !(id_col %in% drop_id)) }
  
  data
  
}

# step 2
StatSfscoperegion <- ggplot2::ggproto(`_class` = "StatSfscoperegion",
                                `_inherit` = ggplot2::Stat,
                                # required_aes = c("fips|county_name"),
                                compute_panel = compute_panel_scope_region,
                               default_aes = ggplot2::aes(label = after_stat(id_col)))


stat_region <- function(
      mapping = NULL,
      data = NULL,
      geom = ggplot2::GeomSf,
      position = "identity",
      na.rm = FALSE,
      show.legend = NA,
      inherit.aes = TRUE,
      crs = "NAD27", # "NAD27", 5070, "WGS84", "NAD83", 4326 , 3857
      ...) {

  c(ggplot2::layer_sf(
              stat = StatSfscoperegion,  # proto object from step 2
              geom = geom,  # inherit other behavior
              data = data,
              mapping = mapping,
              position = position,
              show.legend = show.legend,
              inherit.aes = inherit.aes,
              params = rlang::list2(na.rm = na.rm, ...)
              ),
              
              coord_sf(crs = crs,
                       default_crs = sf::st_crs(crs),
                       datum = crs,
                       default = TRUE)
     )
  }
```




### test it out `stat_county()`

```{r}
source("./R/stat_state.R")
source("./R/stat_muni.R")



library(ggplot2)


br_state |>
  sf::st_drop_geometry() |>
  ggplot() +
  aes(state_name = name_state) +
  stat_state(alpha = .4) + 
  aes(fill = name_region)

br_muni |>
  sf::st_drop_geometry() |>
  ggplot() +
  aes(muni_code = code_muni) +
  stat_muni(alpha = .4, linewidth = .02) + 
  aes(fill = abbrev_state)

```


## Use template to create useful derivitive functions

```{r}
readme2pkg::chunk_variants_to_dir(chunk_name = "geom_region_template",
                                  file_name = "geom_state.R",
                                  replace1 = "region",
                                  replacements1 = "state")


readme2pkg::chunk_variants_to_dir(chunk_name = "geom_region_template",
                                  file_name = "geom_muni.R",
                                  replace1 = "region",
                                  replacements1 = "muni")
```


```{r geom_region_template, eval = F}
geom_region <- stat_region
geom_region_label <- function(...){stat_region(geom = "text",...)}
stamp_region <- function(...){
  stat_region(stamp = T, 
              data = mtcars,
              aes(fill = NULL, color = NULL, label = NULL, 
                  fips = NULL, region_name = NULL), 
              ...)}
stamp_region_label <- function(...){
  stat_region(stamp = T, 
              geom = "text", 
              data = mtcars, 
              aes(fill = NULL, color = NULL,
                  fips = NULL, region_name = NULL), 
              ...)}
```




### try those out

```{r}
source("./R/geom_state.R")
source("./R/geom_muni.R")


br_state |>
  sf::st_drop_geometry() |>
  ggplot() +
  aes(state_name = name_state) +
  geom_state() + 
  geom_state_label(check_overlap = T,
                    color = "grey85") +
  aes(fill = name_region) 

br_muni |>
  sf::st_drop_geometry() |>
  ggplot() +
  aes(muni_code = code_muni) + 
  geom_muni(linewidth = .02) + 
  geom_muni_label(check_overlap = T,
                    color = "grey85", size = 3) +
  aes(fill = abbrev_state)

ggplot() + 
  stamp_state()

last_plot() + 
  stamp_state_label(check_overlap = T)

last_plot() + 
  stamp_state(keep_id = "Amapá", fill = "darkred") +
  stamp_state_label(keep_id = "Amapá", color = "oldlace")

ggplot() + 
  stamp_muni(linewidth = .02)

```


## Use template to write convenience functions for each region

```{r eval= F}
locations <- geo_reference_brasil_state$state_name
locations_snake <- tolower(locations) |> 
  stringr::str_replace_all(" ", "_")

readme2pkg::chunk_variants_to_dir(chunk_name = "stamp_region_location", 
                                  file_name = "stamp_state_location.R",
                                  replace1 = "region",
                                  replacements1 = rep("state", length(locations)),
                              replace2 = "location",
                              replacements2 = locations_snake,
                              replace3 = "Location", 
                              replacements3 = locations)


locations <- geo_reference_brasil_muni$muni_name
locations_snake <- tolower(locations) |> 
  stringr::str_replace_all(" ", "_") |>
  stringr::str_replace_all("'", "_") |>
  stringr::str_replace_all("-", "_")

readme2pkg::chunk_variants_to_dir(chunk_name = "stamp_region_location", 
                                  file_name = "stamp_muni_location.R",
                                  replace1 = "region",
                                  replacements1 = rep("muni", length(locations)),
                              replace2 = "location",
                              replacements2 = locations_snake,
                              replace3 = "Location", 
                              replacements3 = locations)


```


```{r stamp_region_location}
#' Title
#'
#' @param ... 
#'
#' @return
#' @export
#'
#' @examples
stamp_region_location <- function(...){stamp_region(keep_id = "Location", ...)}

#' Title
#'
#' @param ... 
#'
#' @return
#' @export
#'
#' @examples
stamp_region_label_location <- function(...){stamp_region_label(keep_id = "Location", ...)}

```

### Try it out

```{r last_test}
source("./R/stamp_state_location.R")
source("./R/stamp_muni_location.R")


#library(ggbrasil2)

ggplot() +
  aes(state_name = name_state) + 
  stamp_state() + 
  stamp_state_espirito_santo(fill = "green") +
  stamp_state_label_goiás() + 
  stamp_state_label_maranhão() + 
  stamp_state_distrito_federal(fill = "red")

ggplot() +
  stamp_state() +
  stamp_muni_alvarães(fill = "magenta") +
  stamp_muni_água_azul_do_norte(fill = "blue")

```
